---
title: "Linear Regression Experiments"
author: ""
date: "9/24/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = NA, include = TRUE)
set.seed(4857)
```

```{r}
library(cmdstanr)
library(data.table)
library(kableExtra)
library(rjson)
library(gridExtra)
set_cmdstan_path("/home/breck/.cmdstanr/cmdstan-2.27.0")
source(here::here("R","util.R"))
source(here::here("R","sim_configs.R"))
source(here::here("R", "data_configs.R"))
source(here::here("R","modeling_configs.R"))
```
# Running linear regression on Brazil simulations

The 2020 data looks like a candidate for fitting with linear regression. First
we collect the actual data in `r2` and apply a simulated version with `r2_sim`. 


```{r}
r1 <- setup_run_df(seed = 93435, n_pop = 214110287, n_days = 291)
r2 <- data_brazil_2020(r1) #actual 
r2_sim <- sim_Brazil2020(r1)
```

The below plot shows a reasonable emulation of the real data with simulated SIRD
output from the SIRD model `baseline.stan`. 

```{r fig.width = 6}
plot_sim(r2, r2_sim)
```
## Linear Regression

How well does linear regression fit the deaths? We will use simulated data. 

First we have to add configuration for running a Stan model:

```{r}
r3 <- model_stan_linear_reg(r2_sim) #in R/modeling_configs.R
kable(t(trim_for_printing(r3,60)))
```

Code for running `stan/linear_reg.stan`:
```{r}
run_linear_regression <- function(run_df) {
  j <- 0
  while (j < nrow(run_df)) {
    j <- j + 1
    fit <- NA
    if (dir.exists(here::here("output",run_df[j,]$dir_name))) {
      print(paste("Deleting directory", 
                  here::here("output", run_df[j,]$dir_name)))
      unlink(here::here("output",run_df[j,]$dir_name), recursive = TRUE)
    }
    dir.create(here::here("output",run_df[j,]$dir_name))
    stan_data <-
      list(n_days = run_df[j,]$n_days,
           tweets = unlist(run_df[j,]$tweets),
           deaths = unlist(run_df[j,]$d),
           compute_likelihood = run_df[j,]$compute_likelihood,
           use_tweets = run_df[j,]$use_tweets,
           scale = run_df[j,]$scale,
           days_held_out = run_df[j,]$truncate_data,
           debug = 0)
    model <- cmdstan_model(here::here("stan", "linear_reg.stan"))
    write(toJSON(run_df[j,]), here::here("output",run_df[j,]$dir_name, 
                                         "config.json"))
    data = here::here("output",run_df[j,]$dir_name, "data.json")
    write(toJSON(stan_data), data)
    fit <- model$sample(data=stan_data,
                        output_dir = here::here("output/",run_df[j,]$dir_name),
                        parallel_chains = 4,
                        iter_warmup = 1000,
                        iter_sampling = 1000,
                        chains = 4,
                        seed = 4857)
    run_df[j,]$fit = list(fit)
  }
}
```

### Prior predictive check

We start with the traditional prior predictive check. 

```{r}
plot_pred_check <- function(df, fit, n_draws) {
  plot <- ggplot(data = NULL, aes(x = day, y = count))
  plot <- plot_draws(plot = plot, variable = 'pred_deaths', n_draws = n_draws,
                     n_columns = df$n_days, 
                    'blue', fit)
  plot <- graph_observed_data(data_df = df, plot = plot)
  plot <- plot + xlim(0, 330) +
  theme(legend.position = "none") +
  ggtitle(df$description)
}
```

```{r}
r3$use_tweets <- 0
r3$compute_likelihood <- 0
r3$scale <- 0
run_linear_regression(r3)
fit3 <- as_cmdstan_fit(files = list.files(path=here::here("output",r3$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
r4 <- copy_run(r3,"scale")
r4$scale <- 1
run_linear_regression(r4)
fit4 <- as_cmdstan_fit(files = list.files(path=here::here("output",r4$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
```
```{r}

plot3 = plot_pred_check(r3, fit3, 4000) + ylim(0,200000) +
  ggtitle("Prior Predictive Check for Deaths", subtitle = "stan/linear_reg.stan, non-scaled")

plot4 = plot_pred_check(r4, fit4, 4000) + ylim(0,200000) +
  ggtitle("Prior Predictive Check for Deaths", subtitle = "stan/linear_reg.stan, scaled")

grid.arrange(plot3, plot4, ncol=2)
```


```{r}
r5 <- copy_run(r3, 'post-pred-check')
r5$use_tweets <- 1 #don't predict tweets
r5$compute_likelihood <- 1
r5$scale <- 1
run_linear_regression(r5)
fit5 <- as_cmdstan_fit(files = list.files(path=here::here("output",r5$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
plot5 = plot_pred_check(r5, fit5, 40) +
  ggtitle("Posterior Predictive Check for Deaths", subtitle = "stan/linear_reg.stan, scaled")
plot5

```

```{r}
r2_drawn <- sim_jitter_from_sim(r2_sim, 3)
r3 <- model_stan_linear_reg(r2_drawn) #in R/modeling_configs.R
r3$use_tweets = 1
r3$description = paste(r3$description,"tweets")
r4 <- model_stan_linear_reg(r2_drawn)
r4$use_tweets = 0
r4$description = paste(r3$description,"no tweets")
r5 = rbind(r3,r4)

run_linear_regression(r5)


```

## Show fits for data jittered from Brazil data

```{r}
plots = list()
for (i in 1:nrow(r5)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r5[i,]$dir_name), 
                                    pattern = "*.csv$", full.names = TRUE))
  plots[[i]] <- plot_pred_check(r5[i,], fit, 40)
} 
```
```{r fig.height=8}
grid.arrange(grobs=plots, ncol=2)
```



## Fits for data truncated at 140 days, jittered data sets.

```{r message=FALSE, warning=FALSE, include=FALSE}
r6 <- copy(r5)
r6$truncate_data = 140
r6$description = paste(r6$description,"140 trunc")

run_linear_regression(r6)
```
```{r fig.height=8}
plots = list()
for (i in 1:nrow(r6)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r6[i,]$dir_name), 
                                    pattern = "*.csv$", full.names = TRUE))
  plots[[i]] <- plot_pred_check(r6[i,], fit, 40)
} 
grid.arrange(grobs=plots, ncol=2)
```
