---
title: "Simulations Based on Brazil 2020"
author:
  - "Breck Baldwin"
  - "Jose Storopoli"
  - "Conor Rosato"
date: "9/24/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = NA, include = TRUE)
set.seed(4857)
```



```{r}
library(data.table)
library(kableExtra)
source(here::here("R","util.R"))
source(here::here("R","sim_configs.R"))
source(here::here("R", "data_configs.R"))
```

# Simulation runs based on Brazil 2020

Below is code to extract and put Brazil 2020 data into data running infrastructure.
Prints transposed data frame row with a 60 character limit on each entry.
```{r}
r1 <- setup_run_df(seed = 93435, n_pop = 214110287, n_days = 291)
r2 <- data_brazil_2020(r1) #actual 
kable(t(trim_for_printing(r2, 60)))
#trim_for_printing(r2, length=10)
```

## Brazil Data Graphed

Graph Brazil tweets and deaths for 2020.
```{r}
plot <- ggplot(data = NULL, aes(x = day, y = count))
plot <- graph_observed_data(data_df = r2, plot = plot)
plot <- plot + xlim(0, 400) + theme(legend.position = "none")
plot
```
## Simulating Brazil Data 

Above Brazil data was fit with SIRD model from `baseline.stan`. The resulting 
estimates were then used to simulate with a separate SIRD simulation model. 

```{r}
r2_sim <- sim_Brazil2020(r1)
kable(t(trim_for_printing(r2_sim, 60)))
```
## Single simulation of data with actual data plotted as well.
```{r fig.width = 6}
plot_sim(r2, r2_sim)
```

### 8 Simulations with jittered data points

Simulations that are jittered between 0 and 3x patient zero count, +-.1 for for 
other parameters. Transposed values below, all values truncated at 15 characters.

```{r}
r2_drawn <- sim_jitter_from_sim(r2_sim, 7)
kable(t(trim_for_printing(r2_drawn,15)))
```
Plots for simulations, upper left the direct simulation of Brazil 2020.

```{r}
library(gridExtra)
plots = list()
for (i in 1:nrow(r2_drawn)) {
  plots[[i]] <- plot_sim(r2, r2_drawn[i,])
} 
```
```{r fig.height=8}
grid.arrange(grobs=plots, ncol=2)
```