---
title: "Baseline SIRD Runs"
author: "Breck Baldwin"
date: "10/10/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, comment = NA, include = TRUE)
set.seed(4857)
```

```{r}
library(cmdstanr)
library(data.table)
library(kableExtra)
library(rjson)
library(gridExtra)
set_cmdstan_path("/home/breck/.cmdstanr/cmdstan-2.27.0")
source(here::here("R","util.R"))
source(here::here("R","sim_configs.R"))
source(here::here("R", "data_configs.R"))
source(here::here("R","modeling_configs.R"))
```
# Running linear regression on Brazil simulations

The 2020 data looks like a candidate for fitting with linear regression. First
we collect the actual data in `r2` and apply a simulated version with `r2_sim`. 


```{r}
r1 <- setup_run_df(seed = 93435, n_pop = 214110287, n_days = 291)
r2 <- data_brazil_2020(r1) #actual 
r2_sim <- sim_Brazil2020(r1)
```

The below plot shows a reasonable emulation of the real data with simulated SIRD
output from the SIRD model `baseline.stan`. 

```{r fig.width = 6}
plot_sim(r2, r2_sim)
```
## Baseline SIRD 

How well does the SIvRD model fit the deaths? We will use simulated data. 

First we have to add configuration for running a Stan model:

```{r}
r3 <- model_stan_baseline(r2) #in R/modeling_configs.R
kable(t(trim_for_printing(r3,60)))
```

Code for running `stan/SIvRD.stan`:
```{r}
run_SIvRD <- function(run_df) {
  j <- 0
  while (j < nrow(run_df)) {
    j <- j + 1
    fit <- NA
    if (dir.exists(here::here("output",run_df[j,]$dir_name))) {
      print(paste("Deleting directory", 
                  here::here("output", run_df[j,]$dir_name)))
      unlink(here::here("output",run_df[j,]$dir_name), recursive = TRUE)
    }
    dir.create(here::here("output",run_df[j,]$dir_name))
    stan_data <-
      list(n_days = run_df[j,]$n_days,
           sDay1 = run_df[j,]$n_pop - 1,
           iDay1 = 1,
           rDay1 = 0,
           dDay1 = 0,
           center = 0,
           run_twitter = run_df[j,]$use_tweets,
           Npop = run_df[j,]$n_pop,
           tweets = unlist(run_df[j,]$tweets),
           deaths = unlist(run_df[j,]$d),
           compute_likelihood = run_df[j,]$compute_likelihood,
           use_tweets = run_df[j,]$use_tweets,
           run_block_ODE = ifelse(run_df[j,]$ode_solver == 'block', 1, 0),
           run_rk45_ODE = ifelse(run_df[j,]$ode_solver == 'rk45', 1, 0),
           scale = run_df[j,]$scale,
           prior_beta_mean = .3,
           prior_beta_std = .2,
           prior_gamma_mean = .3,
           prior_gamma_std = .2,
           prior_death_prob = .02,
           prior_death_prob_std = .005,
           prior_twitter_lambda = 1.0,
           prior_twitter_std = 1.0,
           I2DandR = 0,
           I2D2R = 1,
           days_held_out = run_df[j,]$truncate_data,
           debug = 0,
           beta_epoch = 30)
    model <- cmdstan_model(here::here("stan", "SIvRD.stan"))
    write(toJSON(run_df[j,]), here::here("output",run_df[j,]$dir_name, 
                                         "config.json"))
    data = here::here("output",run_df[j,]$dir_name, "data.json")
    write(toJSON(stan_data), data)
    fit <- model$sample(data=stan_data,
                        output_dir = here::here("output/",run_df[j,]$dir_name),
                        parallel_chains = 4,
                        iter_warmup = 1000,
                        iter_sampling = 1000,
                        chains = 4,
                        seed = 4857)
    run_df[j,]$fit = list(fit)
  }
}
```

### Prior predictive check

We start with the traditional prior predictive check on actual data.


```{r}
r3$use_tweets <-  0
r3$compute_likelihood <- 0
#run_baseline_SIRD(r3)
#fit3 <- as_cmdstan_fit(files = list.files(path=here::here("output",r3$dir_name), 
#                                         pattern = "*.csv$", full.names = TRUE))
r4 <- copy_run(r3,"prior_pc")
r4$scale <- 0
run_SIvRD(r4)
fit4 <- as_cmdstan_fit(files = list.files(path=here::here("output",r4$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
```
```{r}

#plot3 = plot_pred_check(r3, fit3, 4000) + ylim(0,200000) +
#  ggtitle("Prior Predictive Check for Deaths", subtitle = "stan/linear_reg.stan, non-scaled")

plot4 = plot_pred_check(r4, fit4, 400) + ylim(0,200000) +
  ggtitle("Prior Predictive Check for Deaths", subtitle = "stan/baseline.stan, scaled")
plot4

#grid.arrange(plot3, plot4, ncol=2)
```

Not too sure what to think of the above plot. The predicted deaths are predicted
but the shapes are all wrong. Got the following warnings:

Warning: 2113 of 4000 (53.0%) transitions ended with a divergence.
This may indicate insufficient exploration of the posterior distribution.
Possible remedies include: 
  * Increasing adapt_delta closer to 1 (default is 0.8) 
  * Reparameterizing the model (e.g. using a non-centered parameterization)
  * Using informative or weakly informative prior distributions 

```{r}
fit4
```

The summary appears to cover the fitted parameters used to create the simulated data.

### Prior predictive check on simulated data

How does the check work with simulated SIRD output?


```{r}

r3_sim <- model_stan_baseline(r2_sim)
r3_sim$use_tweets <-  0
r3_sim$compute_likelihood <- 0
r3_sim$scale <- 0
#run_baseline_SIRD(r3)
#fit3 <- as_cmdstan_fit(files = list.files(path=here::here("output",r3$dir_name), 
#                                         pattern = "*.csv$", full.names = TRUE))
r4_sim <- copy_run(r3_sim,"prior_pc")
run_SIvRD(r4_sim)
fit4_sim <- as_cmdstan_fit(files = list.files(path=here::here("output",r4_sim$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))

```


```{r}

plot4_sim = plot_pred_check(r4, fit4_sim, 400) + ylim(0,200000) +
  ggtitle("Prior Predictive Check for Deaths Simulated Data", subtitle = "stan/baseline.stan, scaled")
plot4_sim

#grid.arrange(plot3, plot4, ncol=2)
```

## Posterior predictive check simulated Brazil data

```{r}

r5_sim <- copy_run(r3_sim,"post_pc_Iv")
r5_sim$use_tweets <- 0 
r5_sim$scale <- 1
r5_sim$compute_likelihood <- 1
run_SIvRD(r5_sim)
fit5_sim <- as_cmdstan_fit(files = list.files(path=here::here("output",r5_sim$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))

plot5 = plot_pred_check(r5_sim, fit5_sim, 40) +
  ggtitle("Posterior Predictive Check for Deaths Sim Data", subtitle = "stan/baseline.stan, scaled")
plot5

```

Very tight fit. Lets compare to the actual data:
```{r}

r <- copy_run(r3,"post_pc_real_data_Iv")
r$use_tweets <- 0 
r$scale <- 1
r$compute_likelihood <- 1
run_SIvRD(r)
fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))

plot = plot_pred_check(r, fit, 40) +
  ggtitle("Posterior Predictive Check for Deaths Real Data", subtitle = "stan/baseline.stan, scaled")
plot
```

Less tight a fit, the standard deviation of the estimates is clear and doing a good job of covering
the data. 

Lets add tweets to the predictions

```{r}


r <- copy_run(r3,"post_pc_real_data_tweets")
r$use_tweets <- 1
r$scale <- 1
r$compute_likelihood <- 1
run_SIvRD(r)

```
```{r}
fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))

plot_pred_check(r, fit, 40) +
  ggtitle("Posterior Predictive Check for Deaths Real Data + tweets", subtitle = "stan/baseline.stan, scaled")
plot
```


## Plot Brazil simulation with varied beta 7 day epoch

```{r}
r_model <- model_stan_baseline(r2_sim)
runs = data.frame()
for (i in 1:6) {
  r_copy <- copy_run(r_model,paste0("Iv",i))
  runs <- rbind(runs,r_copy)
}
r <- vary_beta_by_epoch(runs, sd=.1 , epoch=7)
r$use_tweets <- 0
r$scale <- 1
```

```{r}
run_SIvRD(r)
```

```{r fig.height=8}
plots = list()
for (i in 1:nrow(r)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r[i,]$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
  plots[[i]] <- plot_pred_check(r[i,], fit, 40)
} 
grid.arrange(grobs=plots, ncol=2)
```

## Compare recovered betas from simulated betas

```{r fig.height=8}
plots = list()
for (i in 1:nrow(r)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r[i,]$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
  
  #plot_sim <- function(df_data, df_sim) {
  df_sim <- data.frame(day = 1:r[i,]$n_days, 
             beta = unlist(r[i,]$daily_betas),
             sim_or_est = rep(paste("sim", i) ,r[i,]$n_days))
  
  plot <- ggplot(data = NULL, aes(x=day, y=beta))
  plot <- plot_draws(plot = plot, variable = 'beta', n_draws = 20, n_columns=r[i,]$n_days, 
                     color='blue', fit=fit, y_label = 'beta') +
    geom_line(data = df_sim, aes(color = sim_or_est))
  plots[[i]] <- plot
  
}
grid.arrange(grobs=plots, ncol=2)
```
## Plot Brazil simulation with varied beta 30 day epoch

```{r}
r_model <- model_stan_baseline(r2_sim)
r = data.frame()
for (i in 1:6) {
  r_copy <- copy_run(r_model,paste0("Iv",i))
  r_i <- vary_beta_by_epoch(r_copy, sd=.1 , epoch=30)
  r_i$use_tweets <- 0
  r_i$scale <- 1
  #plot <- plot <- ggplot(data = NULL, aes(x=day, y=beta))
  #print(plot_varied_betas(r, plot = plot))
  r <- rbind(runs,r_i)
}


```

```{r}
run_SIvRD(r)
```

```{r fig.height=8}
plots = list()
for (i in 1:nrow(r)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r[i,]$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
  plots[[i]] <- plot_pred_check(r[i,], fit, 40)
} 
grid.arrange(grobs=plots, ncol=2)
```

## Compare recovered betas from simulated betas

```{r fig.height=8}
plot_varied_betas <- function(r, plot) {
   df_sim <- data.frame(day = 1:r$n_days, 
               beta = unlist(r$daily_betas),
               sim = rep(paste("sim", i) ,r$n_days))
   return (plot + geom_line(data = df_sim, aes(color = sim)))
}

plot_varied_betas_and_sims <- function(r, n_draws){
  plots = list()
  for (i in 1:nrow(r)) {
    #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
    fit <- as_cmdstan_fit(files =
                            list.files(path=here::here("output",r[i,]$dir_name), 
                                       pattern = "*.csv$", full.names = TRUE))
    
    plot <- ggplot(data = NULL, aes(x=day, y=beta))
    plot <- plot_predictions(plot, prediction_label = 'beta', fit = fit,
                             show_ribbon = TRUE, y_label = 'beta') 
    #plot <- plot_draws(plot = plot, variable = 'beta', n_draws = n_draws, 
    #                   n_columns=r[i,]$n_days, color='blue', fit=fit, 
    #                   y_label = 'beta')
      
    plots[[i]] <- plot_varied_betas(r[i,], plot = plot)
    
  }
  return(plots)
}
plots <- plot_varied_betas_and_sims(r, n_draws = 20)
grid.arrange(grobs = plots, ncol = 2)
```

what does beta plot look like for one beta epoch?
graph mean/ribbon as well
match epoch in model to sim epoch



## Show fits for data jittered from Brazil data


```{r}
r2_drawn <- sim_jitter_from_sim(r2_sim, 3)
rj <- model_stan_baseline(r2_drawn) #in R/modeling_configs.R
rj$use_tweets <- 1
rj$compute_likelihood <- 1
rj$scale <- 1
run_SIvRD(rj)

```


```{r}
plots = list()
for (i in 1:nrow(rj)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",rj[i,]$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))
  plots[[i]] <- plot_pred_check(rj[i,], fit, 40)
} 
```
```{r fig.height=8}
grid.arrange(grobs=plots, ncol=2)
```



## Fits for data truncated at 140 days, jittered data sets.

```{r message=FALSE, warning=FALSE, include=FALSE}
r6 <- copy(rj)
r6$truncate_data = 140
r6$description = paste(r6$description,"140 trunc Iv")

run_SIvRD(r6)
```
```{r fig.height=8}
plots = list()
for (i in 1:nrow(r6)) {
  #plots[[i]] <- plot_sim(r2, r2_drawn[i,])
  fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r6[i,]$dir_name), 
                                    pattern = "*.csv$", full.names = TRUE))
  plots[[i]] <- plot_pred_check(r6[i,], fit, 40)
} 
grid.arrange(grobs=plots, ncol=2)
```

 
```{r}


r <- copy_run(r3,"post_pc_real_data_tweets_trunc")
r$use_tweets <- 1
r$scale <- 1
r$compute_likelihood <- 1
r$truncate_data = 140
r$description = paste(r$description,"140 trunc")
run_baseline_SIRD(r)

```
```{r}
fit <- as_cmdstan_fit(files = list.files(path=here::here("output",r$dir_name), 
                                         pattern = "*.csv$", full.names = TRUE))

plot_pred_check(r, fit, 40) +
  ggtitle("Posterior Predictive Check for Deaths Real Data + tweets", subtitle = "stan/baseline.stan, scaled")
plot
```
